The problem and promise of scale in multilayer networks

### 2020-01-24 paper/manuscript.Rmd
small grammar fixes

In long-lived species, social phenotypes can change over their lifetimes in ways that can affect fitness (Berger et al., 2015; Brent et al., 2017). These findings highlight the necessity to carefully consider the temporal scale that is meaningful for the proposed hypotheses, especially for understanding the relationship between an individuals social phenotypes and fitness. Further complicating the issue, social phenotypes can change over stages of ontogeny, although it is important to note that individuals progress through development at different rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This variation can make defining appropriate temporal scales for testing the fitness implications of an individual’s social phenotype over different stages of ontogeny even more difficult. However, considering fine temporal scales may provide insight across development stages that may otherwise be missed (CITE-myself??). Multilayer networks provide a framework to test hypotheses linking social phenotypes and fitness across temporal scales. Because of the non-independent nature of social interactions and associations, especially over time, it is hard to test hypotheses about social dynamics with a suitable null control (Farine, 2017; Proskurnikov & Tempo, 2017). Using the framework of multilayer networks, making layers of different temporal periods provides a way to make appropriate null models to test questions about the dynamics of an individual’s social network position over multiple aggregated time periods.

Social scale is another major consideration for social network analysis, especially in the context of different types of interactions and associations in the same analysis (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine, 2015). It is a common trope that individuals must be associating to interact and most studies therefore assume that proximity is proxy for interacting (Farine, 2015). For instance, baboons need to be in proximity in order to groom each other, but this trope highlights human bias toward interactions that can be readily observed in the field. Animals have certain interactions that do not require proximity such as long-distance vocalization and olfactory signals (Carter et al., 2015). With the advancement of bio-logging technology, our ability to record social interactions that do not require proximity, particularly long-distance communication networks with devices like microphone arrays (Snijders & Naguib, 2017). The multilayer network framework further enables researchers to examine the relationship between social phenotypes and fitness across social scales and to test how good a social proximity as a proxy for social interactions in a holistic way.

@@

Multilayer networks further provide unique challenges and opportunities for understanding the social phenotypes and environments of individuals and their fitness outcomes. Picking the proper temporal scale is important but a challenge because all layers in the multilayer network should reflect the same time period, and the period should be long enough for each layer to have enough data for robust networks (Farine, 2017). These considerations may be limiting factors when answering questions that require timeframes that are relevant to different individuals. On the other hand, multilayer network metrics provide an opportunity to simplify statistical models. Interactions that happen meters apart, proximity associations, and long-distance interactions can be encapsulated in a single multilayer network and a single metric can be used to describe the various social scales of an individual’s network position. This multilevel network metric could be used to assess the relationship between an individuals social phenotype and fitness. It would also be a meaningful simplification of many network metrics that could help simplify statistical analyses and deal with the frequent correlations between social network metrics in different types of networks (Castles et al., 2014).

>>>>

In long-lived species, social phenotypes can change over their lifetimes in ways that can affect fitness (Berger et al., 2015; Brent et al., 2017). These findings highlight the necessity to carefully consider the temporal scale that is meaningful for the proposed hypotheses, especially for understanding the relationship between an individuals social phenotype and fitness. Further complicating the issue, social phenotypes can change over stages of ontogeny, although it is important to note that individuals progress through development at different rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This variation can make defining appropriate temporal scales for testing the fitness implications of an individual’s social phenotype over different stages of ontogeny even more difficult. However, considering fine temporal scales may provide insight across development stages that may otherwise be missed (CITE-myself??). Multilayer networks provide a framework to test hypotheses linking social phenotypes and fitness across temporal scales. Because of the non-independent nature of social interactions and associations, especially over time, it is hard to test hypotheses about social dynamics with a suitable null control (Farine, 2017; Proskurnikov & Tempo, 2017). Using the framework of multilayer networks, making layers of different temporal periods provides a way to make appropriate null models to test questions about the dynamics of an individual’s social network position over multiple aggregated time periods.

Social scale is another major consideration for social network analysis, especially in the context of different types of interactions and associations in the same analysis (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine, 2015). It is a common trope that individuals must be associating to interact and most studies therefore assume that proximity is proxy for interacting (Farine, 2015). For instance, baboons need to be in proximity in order to groom each other, but this trope highlights human bias toward interactions that can be readily observed in the field. Animals have certain interactions that do not require proximity such as long-distance vocalization and olfactory signals (Carter et al., 2015). With the advancement of bio-logging technology, we have the ability to record social interactions that do not require proximity, particularly long-distance communication networks with devices like microphone arrays (Snijders & Naguib, 2017). The multilayer network framework further enables researchers to examine the relationship between social phenotypes and fitness across social scales and to test how good social proximity is as a proxy for social interactions in a holistic way.

@@

Multilayer networks further provide unique challenges and opportunities for understanding the social phenotypes and environments of individuals and their fitness outcomes. Picking the proper temporal scale is important but a challenge because all layers in the multilayer network should reflect the same time period, and the period should be long enough for each layer to have enough data for robust networks (Farine, 2017). These considerations may be limiting factors when answering questions that require timeframes that are relevant to different individuals. On the other hand, multilayer network metrics provide an opportunity to simplify statistical models. Interactions that happen meters apart, proximity associations, and long-distance interactions can be encapsulated in a single multilayer network, and a single metric can be used to describe the various social scales of an individual’s network position. This multilevel network metric could be used to assess the relationship between an individual's social phenotype and fitness. It would also be a meaningful simplification of many network metrics that could help simplify statistical analyses and deal with the frequent correlations between social network metrics in different types of networks (Castles et al., 2014).
